\chapter*{Acknowledgments}

\setheader{Acknowledgments}

The last four years of work have been marked by the help and support of many
friends and colleagues to whom I want to express my most sincere gratitude.
Most of the work in this thesis wouldn't have been possible without them. 

Prof. Michel Orrit, my supervisor, deserves a big part of the credit for what
has been done in this last four years. His support and guidance where crucial
for the completion of this thesis. I learned a lot in these past years thanks to
his insightful knowledge over a broad range of topics. 

At the beginning of my research, dr. Pedro Navarro and dr. Saumyakhanti
Khatua, two fellow members of the group, were of great importance both for
scientific discussion and for me to integrate into the group. I am also
grateful to the members of the project in which my PhD was framed, including dr.
Aimee Boyle for the majority of the nanorod samples employed throughout the
thesis, Veer Keizer for the help preparing the samples with cells that lead
to chapter 3 and Sara Carozza for several insightful discussions. 

I am also very grateful to dr. Mart\'{i}n Caldarola, with whom I have shared
many hours in the same dark, cold lab. Long discussions with him triggered different
approaches and ideas, some of which are reflected in this thesis. He is not only
a colleague but also a good friend, whose support has extended beyond the walls
of the university.

I am indebted to Ferry Kruidenberg whose bachelor project gave the initial
push to the software to control the setup. I am also indebted to Irina Komen,
whose MSc project was part of a very challenging idea that finally didn't find
its way into this thesis. 

The MoNOS group has provided the perfect setting for conducting research, and I
am grateful to all the members, specially to Henri\"{e}tte van Leeuwen for
taking care of so much paperwork. 

Of course nothing of what happened during these last four years would have
been possible without the endless love and support from my family, who was
always present.